Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems\thanks{This research was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix

In our research we took weight coefficients as nodes of a uniform grid in the polyhedron (\ref{eq:06}). For assessment of the quality of Pareto set approximation obtained after solving problems (\ref{eq:08}) with mentioned weights, the HyperVolume index (HV-index) \cite{Evtushenko2014,Gergel2018} was applied as traditional criterion of approximation accuracy.

@@

In general case, the functions satisfying the Lipschitz condition are multiextremal, therefore, the convolution (\ref{eq:05}) will be multiextremal and Lipschitzian as well. It means that for solving the problem (\ref{eq:08}) it is required to apply methods of nonlinear programming which provides finding the global optimum. There are many global optimization algorithms designed under different assumptions about properties of objective function; relevant references can be found, for instance, in monographs \cite{Pardalos2017,Marler2009,Sergeyev2013}. Some of them are theoretically substantiated and guarantee the convergence to global solution; the others are empirical, for example, nature-inspired methods that do not ensure finding the global optimum.

@@

Existence of Peano mappings can be applied to solving the optimization problems in the following way. Let there be a $n$-dimensional continuous function $g(y)$ defined in the domain (\ref{eq:01}) and $y(x)$ is a Peano-type curve. Then the continuity of $g(y)$ and $y(x)$ allows one to assert that
\begin{equation}
    \min_{y\in H} g(y) = \min_{x\in[0,1]} g(y(x)),
\end{equation}
i.e. instead of minimization of the multidimensional function $g(y)$ in the domain (\ref{eq:01}) one can solve the minimization problem for the one-dimensional function $g(y(x))$ over the interval $[0,1]$.

@@

For this goal, it is useful to organize a special matrix of a posteriori information which will store results of all conducted trials, for example, in the form
\begin{equation}
    \label{eq:18}
A^i_p =
\begin{pmatrix}
x^1 & x^2 & \dots & x^p \\
z^{i,1} & z^{i,2} & \dots & z^{i,p} \\
\omega_1^{i,1} & \omega_1^{i,2} & \dots & \omega_1^{i,p} \\
\dots & \dots & \dots & \dots \\
\omega_k^{i,1} & \omega_k^{i,2} & \dots & \omega_k^{i,p}
\end{pmatrix}
\end{equation}
after implementing $p$ trials at points $x^1, x^2, \dots, x^p$ when minimizing a function $f'(x)$ from (\ref{eq:15}).

@@

\textbf{Rule 3.} In the set of trial coordinates $x_j, 1 \leq j \leq p$, non-dominated points are selected according the rule of Definition 2 and all such points are considered as points belonging to current approximation of the Pareto set. All columns of the information matrix $A^i_p$ are marked up with labels 0 or 1 where 1 corresponds to non-dominated points and 0 to dominated ones.

@@

\textbf{Rule 8.} Among all subintervals, the subinterval $(x_{j-1}, x_j), 1 \leq j \leq p-1$, with the highest characteristic is chosen:
\begin{equation}
    \label{eq:25}
    B(l) = \max_{1 \leq j \leq p-1} {B(j)}.
\end{equation}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{fig2}
\caption{The level curves for test function.} \label{fig:2}
\end{figure}

>>>>

In our research we took weight coefficients as nodes of a uniform grid in the polyhedron (\ref{eq:06}). For assessment of the quality of Pareto set approximation obtained after solving problems (\ref{eq:08}) with mentioned weights, the HyperVolume index (HV-index) \cite{Evtushenko2014,Gergel2018} was applied as a traditional criterion of the approximation accuracy.

@@

In general case, the functions satisfying the Lipschitz condition are multiextremal, therefore, the convolution (\ref{eq:05}) will be multiextremal and Lipschitzian as well. It means that for solving the problem (\ref{eq:08}) it is required to apply methods of nonlinear programming which provide finding the global optimum. There are many global optimization algorithms designed under different assumptions about properties of objective function; relevant references can be found, for instance, in monographs \cite{Pardalos2017,Marler2009,Sergeyev2013}. Some of them are theoretically substantiated and guarantee the convergence to global solution; the others are empirical, for example, nature-inspired methods that do not ensure finding the global optimum.

@@

Peano mappings can be applied to solving the optimization problems in the following way. Let there be a $n$-dimensional continuous function $g(y)$ defined in the domain (\ref{eq:01}) and $y(x)$ is a Peano-type curve. Then the continuity of $g(y)$ and $y(x)$ allows one to assert that
\begin{equation}
    \min_{y\in H} g(y) = \min_{x\in[0,1]} g(y(x)),
\end{equation}
i.e. instead of minimization of the multidimensional function $g(y)$ in the domain (\ref{eq:01}) one can solve the minimization problem for the one-dimensional function $g(y(x))$ over the interval $[0,1]$.

@@

For this goal, it is useful to organize a special matrix of a posteriori information which will store results of all conducted trials, for example, in the form
\begin{equation}
    \label{eq:18}
A^i_p =
\begin{pmatrix}
x^1 & x^2 & \dots & x^p \\
z^{i,1} & z^{i,2} & \dots & z^{i,p} \\
\omega_1^{i,1} & \omega_1^{i,2} & \dots & \omega_1^{i,p} \\
\dots & \dots & \dots & \dots \\
\omega_k^{i,1} & \omega_k^{i,2} & \dots & \omega_k^{i,p}
\end{pmatrix}
\end{equation}
after implementing $p$ trials at points $x^1, x^2, \dots, x^p$ when minimizing a function $f^i(x)$ from (\ref{eq:15}).

@@

\textbf{Rule 3.} In the set of trial coordinates $x_j, 1 \leq j \leq p$, non-dominated points are selected according to the rule of Definition 2 and all such points are considered as points belonging to current approximation of the Pareto set. All columns of the information matrix $A^i_p$ are marked up with labels 0 or 1 where 1 corresponds to non-dominated points and 0 to dominated ones.

@@

\textbf{Rule 8.} Among all subintervals, the subinterval $(x_{l-1}, x_l), 1 \leq l \leq p-1$, with the highest characteristic is chosen:
\begin{equation}
    \label{eq:25}
    B(l) = \max_{1 \leq j \leq p-1} {B(j)}.
\end{equation}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig2}
\caption{The level curves for test function.} \label{fig:2}
\end{figure}
